Updates a Procrustes and a PCA

### 2025-05-01 Analysis/17-update_procrustes_and_PCA.Rmd
Update 17-update_procrustes_and_PCA.Rmd

TODO: TRY CENTERING THE DATA BEFORE DOING THE PCA

@@

## Test with pipeline
updated_pca <- update.gpa.pca(proc_orig, new_spec)*-1 # TODO: fix the -1
plot(proc_orig$PCscores[,c(1,2)])
points(updated_pca[,c(1,2)], col = "grey", pch = 19)
points(updated_pca[c(81:85),c(1,2)], col = "orange", pch = 19)
points(updated_pca[c(81:83),c(1,2)], col = "blue", pch = 21)

@@

# No pipeline
# Get the CPA
original_data <- whole_gpa
new_landmarks <- new_individuals
original_pca <- prcomp(landvR::array.to(original_data$rotated, to = "matrix")) # SOLUTION IS SOMEWHERE BETWEEN THIS AND THE PRCOMPFAST
# original_pca <- prcompfast(original_data$tan) # SOLUTION IS SOMEWHERE BETWEEN THIS AND THE PRCOMPFAST
## Project the new specimens on the original GPA
new_data <- Morpho::align2procSym(original_data, new_landmarks)
## Project the new GPAed specimens onto the PCA
new_pca <- stats::predict(original_pca, newdata = array.to(new_data, to = "matrix"))
## Combine the new and old matrices for the output PCA
updated_pca <- rbind(original_pca$x, new_pca) *-1

>>>>

@@

## Test with pipeline
updated_pca <- update.gpa.pca(proc_orig, new_spec)#*-1 # TODO: fix the -1
plot(proc_orig$PCscores[,c(1,2)])
points(updated_pca[,c(1,2)], col = "grey", pch = 19)
points(updated_pca[c(81:85),c(1,2)], col = "orange", pch = 19)
points(updated_pca[c(81:83),c(1,2)], col = "blue", pch = 21)

@@

## From Morpho:
eigenPCA <- function(data,tol=.Machine$double.eps) {
    center1 <- colMeans(data,2,mean)
    datascale <- scale(data,2,scale=FALSE)
    eigData <- eigen(cov(datascale),symmetric=TRUE)
    eigData$values <- eigData$values[which(eigData$values > tol)] 
    sdev <- sqrt(eigData$values)
    rotation <- eigData$vectors[,1:length(sdev)]
    x <- t(t(rotation)%*%t(datascale))
    return(list(x=x,rotation=rotation,sdev=sdev,center=center1))
}

# No pipeline
# Get the CPA
original_data <- whole_gpa
new_landmarks <- new_individuals
original_pca <- prcomp(landvR::array.to(original_data$rotated, to = "matrix")) # SOLUTION IS SOMEWHERE BETWEEN THIS AND THE PRCOMPFAST
original_pca <- prcompfast(original_data$tan*-1) # SOLUTION IS SOMEWHERE BETWEEN THIS AND THE PRCOMPFAST
original_pca <- eigenPCA(original_data$tan)

    center1 <- colMeans(data,2,mean)
    datascale <- scale(data,2,scale=FALSE)
    eigData <- eigen(cov(datascale),symmetric=TRUE)

landvR::array.to(original_data$rotated, to = "matrix")

## Project the new specimens on the original GPA
new_data <- Morpho::align2procSym(original_data, new_landmarks)
## Project the new GPAed specimens onto the PCA
new_pca <- stats::predict(original_pca, newdata = array.to(new_data, to = "matrix"))
## Combine the new and old matrices for the output PCA
updated_pca <- rbind(original_pca$x, new_pca) *-1
